Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-08 paper/paper.tex
#22 add project description to intro

In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.

\cite{levine2013, levine2021} argue that social evolution is driven by conflict. Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English. Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia. In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony. In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p. 439). strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.

>>>>

In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development. If the effect is positive, then selection of "better" states through conflict-driven evolution may partially explain the ascendancy of Europe.

I do this by looking at the development of cities after they switch states. In a paper from which I draw heavily, \cite{schoenholzer2022} find that switching states affects population negatively in the short run but positively in the long run. I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct. 
Using these concepts I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B,

absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires

the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking,

If a city is absorbed via conquest,

Past treatments of this topic have suffered from a lack of available historical data. To empirically tackle it one needs at least two things: the detailed territorial history of each city, and some city-level proxy for development, both reaching far enough into the past and ideally recorded in short intervals.

\cite{diamond1997} and \cite{landes2006} present narrative arguments for the economic benefits of state competition but stop short of formal analysis. 
More recently \cite{schoenholzer2022}, from whom I borrow extensively in this paper, use a dataset with a bigger geographical scope but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years. They are still able to estimate short-run (same-period) and long-run (next two centuries) effects of switching states. 
The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains

Rather than m

\cite{diamond1997} asserts that the geographical remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled innovation in medieval China would not have survived long in a more fragmented state system.

\cite{levine2013, levine2021} provide the theoretical foundation for conflict-driven selection, as well as compelling evidence for it. 
They argue that social evolution is driven by conflict. Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English. Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia. In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony. In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p. 439). strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.
